---
title: "follow up QC for MCS"
author: "Biyao Wang"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Index

The current document describes:

1. Genetic quality control.
2. Principal component analysis.
3. Identify and extract complete trios.

# Qualitry control
## Preliminary QC in plink 

```
DIR="/lustre/scratch/scratch/ucjues9/mcs/genotype/output/"

module load plink/1.90b3.40

plink --bfile ${DIR}mcs_imputed_allchr \
      --mind 0.05  --geno 0.05 --maf 0.01 --hwe 0.000001 \
      --extract /home/ucjues9/Scratch/mcs/genotype/rsq/70.rsq \
      --make-bed --out ${DIR}mcs_imputed_allchr_minQC
```

## Follow-up quality control

```{r, eval=FALSE}
cd /lustre/projects/CMAP/data/MCS/genotype/plink

wc -l mcs_imputed_allchr.bim         # 38909200 SNPs
wc -l mcs_imputed_allchr.fam         # 21349 individuals

wc -l mcs_imputed_allchr_minQC.bim   # 6948515 SNPs
wc -l mcs_imputed_allchr_minQC.fam   # 21349 individuals
```

### Update FID and IID 

Create the update id file with the following four fields:

- Old family ID
- Old within-family ID
- New family ID
- New within-family ID

```{r engine = 'bash', eval = FALSE}

cd /lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/output
head mcs_imputed_allchr.fam
awk '{print $1}' mcs_imputed_allchr.fam > fam_onlyIID_raw.txt  # save only the IID column
head fam_onlyIID_raw.txt
cut -d_ -f1,2 fam_onlyIID_raw.txt > fam_onlyIID_new.txt  # remove everything after the second "_"
head fam_onlyIID_new.txt
paste fam_onlyIID_raw.txt fam_onlyIID_new.txt > fam_onlyIID_rawnew.txt 
head fam_onlyIID_rawnew.txt 

module unload compilers mpi
module load r/recommended   
R
```

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}
library(tidyverse)
library(haven)

setwd("/lustre/projects/CMAP/data/MCS/phenotype/MDAC-2020-0016-05A-PINGAULT")
df <- read_spss("MDAC-2020-0016-05A-PINGAULT_unifed_ega_2021_12_13.sav") 
head(df)
fam_pheno <- df %>%
  select(FID_pheno = 2, IID_pheno = 1) 
head(fam_pheno)  

setwd("/lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/output")
fam_geno <- read_tsv("fam_onlyIID_rawnew.txt", col_names = F)
head(fam_geno)
colnames(fam_geno) <- c("IID_geno_raw", "IID_geno_new")
head(fam_geno)

fam_geno$FID_geno_new <- fam_pheno$FID_pheno[match(fam_geno$IID_geno_new, fam_pheno$IID_pheno)]
head(fam_geno)

fam_update <- fam_geno[,c(1,1,3,2)]
head(fam_update)
colnames(fam_update) <- c("FID_old", "IID_old", "FID_new", "IID_new")
head(fam_update)

write.table(fam_update, file=paste0("fam_update.txt"),col.names = T, row.names = F, quote=F, sep = "\t")
q(save="no")
```

```{r engine = 'bash', eval = FALSE}

echo '
#!/bin/bash -l
#$ -l h_rt=0:30:00
#$ -l mem=5G
#$ -l tmpfs=15G
#$ -N updateid
#$ -pe smp 16
#$ -wd /lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/log

cd $TMPDIR

cd /lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/output

module load plink/1.90b3.40

plink --bfile mcs_imputed_allchr_minQC \
      --update-ids fam_update.txt \
      --make-bed \
      --out mcs_imputed_allchr_minQC_updateID
' > updateid.sh
qsub updateid.sh

```

### Pruning

```{r engine = 'bash', eval = FALSE}

module load plink/1.90b3.40

plink --bfile ${DIR}/mcs_imputed_allchr \
      --indep-pairwise 250 50 0.1 \
      --out ${DIR}/mcs_imputed_allchr_pruned


```

### Heterozygosity QC

```{r engine = 'bash', eval = FALSE}
module load plink/1.90b3.40

plink --bfile ${DIR}/mcs_imputed_allchr_minQC_updateID \
      --extract ${DIR}/mcs_imputed_allchr_pruned.prune.in \
      --het \
      --out ${DIR}/mcs_imputed_allchr_minQC_updateID_checkhet

```

### List heterozygosity rate outliers

```{r engine = 'bash', eval = FALSE}

module unload compilers mpi
module load r/recommended   
cd /lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/output
R 
```

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}
library(foreign)

het <- read.table("mcs_imputed_allchr_minQC_updateID_checkhet.het", head=TRUE)
het$HET_RATE = (het$"N.NM." - het$"O.HOM.")/het$"N.NM."
het_fail = subset(het, (het$HET_RATE < mean(het$HET_RATE)-3*sd(het$HET_RATE)) | (het$HET_RATE > mean(het$HET_RATE)+3*sd(het$HET_RATE)))
het_fail$HET_DST = (het_fail$HET_RATE-mean(het$HET_RATE))/sd(het$HET_RATE)
write.table(het_fail, "fail-het-qc.txt", row.names=FALSE)

q(save = "no")
```

```{r engine = 'bash', eval = FALSE}
head fail-het-qc.txt
wc -l fail-het-qc.txt  # 547 outliers + 1 header
sed 's/"// g' fail-het-qc.txt | awk '{print $1, $2}' | tail -n +2 > het_fail_list.txt
head het_fail_list.txt
wc -l het_fail_list.txt
```

### Remove heterozygosity rate outliers

```{r engine = 'bash', eval = FALSE}

module load plink/1.90b3.40

plink --bfile mcs_imputed_allchr_minQC_updateID \
      --remove het_fail_list.txt \
      --make-bed \
      --out mcs_imputed_allchr_minQC_updateID_hetQC

```

### Relatedness QC
Check in phenotypic data.

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}
library(tidyverse)
library(glue)
library(haven)

fam_pheno <- read_spss("MDAC-2020-0016-05A-PINGAULT_unifed_ega_2021_12_13.sav") 
head(fam_pheno)
df <- fam_pheno %>%
  select(FID = 2, IID = 1, MFC) 
head(df)  

df %>%
  count(FID, MFC) %>%
  count(MFC, n, name = "total") 
# more than two individual with the same role in a family
# 151 children
# 10 fathers
# 21 mothers

```

## separate children and parents

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}

library(tidyverse)
library(haven)

df <- read_spss("MDAC-2020-0016-05A-PINGAULT_unifed_ega_2021_12_13.sav") 

fam_pheno <- df %>%
  select(IID = 1, MFC) 

fam_geno <- read.table("mcs_imputed_allchr_minQC_updateID.fam", sep = " " , header = F)

colnames(fam_geno) <- c("FID", "IID", "X1", "X2", "X3", "X4")

fam_geno$MFC <- fam_pheno$MFC[match(fam_geno$IID, fam_pheno$IID)]

child <- fam_geno %>%
  filter(MFC == "C") 

parent <- fam_geno %>%
  filter(MFC != "C") 
  
write.table(child[,-7], file=paste0("children.txt"),col.names = F, row.names = F, quote=F, sep = "\t")
write.table(parent[,-7], file=paste0("parents.txt"),col.names = F, row.names = F, quote=F, sep = "\t")

```

```{r engine = 'bash', eval = FALSE}
module load plink/1.90b3.40

plink --bfile ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC \
      --keep ${DIR}/children.txt \
      --make-bed \
      --out ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_children
      
plink --bfile ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC \
      --keep ${DIR}/parents.txt \
      --make-bed \
      --out ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_parents

```

### Get a list of unrelated children and parents 

```{r, eval=FALSE}

# Check relatedness between children
/lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/king  -b ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_children.bed \
      --related \
      --degree 3 \
      --cpus 36 \
      --prefix ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_children_relatedness

# Get a list of unrelated children    
/lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/king  -b ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_children.bed \
      --unrelated \
      --degree 3 \
      --cpus 36 \
      --prefix ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_children_3rd_

wc -l output/mcs_imputed_allchr_minQC_updateID_hetQC_children_3rd_unrelated.txt              # 7822
wc -l output/mcs_imputed_allchr_minQC_updateID_hetQC_children_3rd_unrelated_toberemoved.txt  # 164

# Check relatedness between parents
/lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/king  -b ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_parents.bed \
      --related \
      --degree 3 \
      --prefix ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_parents_relatedness
                     
# Get a list of unrelated parents
/lustre/projects/CMAP/data/MCS/genotype/imputed_QCed/king  -b ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_parents.bed \
      --unrelated \
      --degree 3 \
      --cpus 36 \
      --prefix ${DIR}/mcs_imputed_allchr_minQC_updateID_hetQC_parents_3rd_

wc -l output/mcs_imputed_allchr_minQC_updateID_hetQC_parents_3rd_unrelated.txt              # 12458
wc -l output/mcs_imputed_allchr_minQC_updateID_hetQC_parents_3rd_unrelated_toberemoved.txt  # 241

```

### Remove related individuals

```{r engine = 'bash', eval = FALSE}
cat mcs_imputed_allchr_minQC_updateID_hetQC_children_3rd_unrelated.txt mcs_imputed_allchr_minQC_updateID_hetQC_parents_3rd_unrelated.txt > mcs_imputed_allchr_minQC_updateID_hetQC_unrelated.txt

wc -l mcs_imputed_allchr_minQC_updateID_hetQC_unrelated.txt # 20280

module load plink/1.90b3.40

plink --bfile mcs_imputed_allchr_minQC_updateID_hetQC \
      --keep mcs_imputed_allchr_minQC_updateID_hetQC_unrelated.txt \
      --make-bed \
      --out mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC

```

# Principal Component Analysis

```{r engine = 'bash', eval = FALSE}

module load plink/1.90b3.40

plink --bfile mcs_imputed_allchr_minQC_updateID_hetQC \
      --extract mcs_imputed_allchr_pruned.prune.in \
      --pca \
      --out mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC

# These are the eigenvalues explained by each principal component, by default, plink estimates 20
head mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC.eigenvec

# These are the actual 20 pca for each individual in the data set
```


```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}

# import PCs
PCs = read.table("output/mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC.eigenvec", header=FALSE)
head(PCs)
colnames(PCs) = c("FID", "IID", "PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14", "PC15", "PC16", "PC17", "PC18", "PC19", "PC20")

# export PCs
write.table(PCs, file="output/mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC_PCs.csv",col.names = T, row.names = F, quote=F, sep = "\t")
save(PCs, file = "output/mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC_PCs.RData")
```

# Get complete trios

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}

# Load libraries
library(tidyverse)
library(haven)

# Load data
fam_pheno <- read_spss("MDAC-2020-0016-05A-PINGAULT_unifed_ega_2021_12_13.sav") 

df_pheno <- fam_pheno %>%
  select(FID = 2, IID = 1, MFC) 

fam_geno <- read.table("mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC.fam", sep = " " , header = F)
colnames(fam_geno) <- c("FID", "IID", "PID","MID","Sex","Pheno")

df_geno <- fam_geno %>%
  select(FID = 1, IID = 2) 

df_geno$MFC <- df_pheno$MFC[match(df_geno$IID, df_pheno$IID)]

# extract children, mothers and fathers 
children <- subset(df_geno,MFC=="C")
mothers <- subset(df_geno,MFC=="M")
fathers <- subset(df_geno,MFC=="F")

# family ID of trios
commonFID <- Reduce(intersect, list(children$FID, mothers$FID, fathers$FID))
length(commonFID)    # 3228

# extract trios
trios <- fam_geno[fam_geno$FID %in% commonFID,]

# save trios fam file
write.table(trios, file="mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC_completetrios.txt", row=F, col=F, qu=F)
```

## Extract unrelated complete trios

```{r engine = 'bash', eval = FALSE}

module load plink/1.90b3.40

plink --bfile mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC \
      --keep mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC_completetrios.txt \
      --make-bed \
      --out mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC_trios

wc -l output/mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC_trios.fam

```
