---
title: "Data preparation"
author: "Jose J. Morosoli"
date: "2025-01-21"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

# Index

The current document describes:

1. Computing new variables and creating data sets for analysis.
2. Standardising variables.
3. Creating clinical threshold variable.


## New variables and data sets.

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}
# Load packages
library(foreign)

# Create an object to store the file path
file_path <- "path/to/your/file/"

# Load files Files here were stored locally but are all available on the cluster. 
# PGS were generated previously using LDpred2.

# Download files containing participant phenotypic information.
# CM-answered questionnaires and derived variables
CMstructure <- read.spss(paste(file_path,"GENDAC_PINGAULT_mcs_cm_structure_2021_05_12.sav", sep = ""), to.data.frame=TRUE)

# Parent answered questionnaires
parentstructure <- read.spss(paste(file_path,"GENDAC_PINGAULT_mcs_parent_structure_2021_05_12.sav", sep = ""),  to.data.frame=TRUE)

# Parent on behalf of CM answered questionnaires
parentCMstructure <- read.spss((paste(file_path,"GENDAC_PINGAULT_mcs_parent_cm_structure_2021_05_12.sav", sep = "", to.data.frame=TRUE)

# Load PGS data from LDpred2 analysis 
EApgs <- read.table((paste(file_path,"EA3_no23.csv.gz_pred_auto.txt", sep = ""), header=TRUE)
CPpgs <- read.table((paste(file_path,"CP_all.csv.gz_pred_auto.txt", sep = ""), header=TRUE)
NCPpgs <- read.table((paste(file_path,"NonCog_no23.csv.gz_pred_auto.txt", sep = ""), header = TRUE)

```

### Create dataframes 

Note: Including phenotypic (mental health, demographics etc) and genotypic (PGS) information per trait (EA, CP, NCP).

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}

{
  library(tidyverse)
  library(foreign)
  library(dplyr)
  
  # Load data frame used to match IDs
  matches <- read.spss(paste(file_path,"MDAC-2020-0016-05A-PINGAULT_unifed_ega_2021_12_13.sav", sep = ""), to.data.frame=TRUE)
  
  # Create data frame for standardised SDQ scores
  SDQ <- subset(CMstructure, select=c(Pingault_ID, BEBDTOT, CEBDTOT, DDDEBDTOT, FEBDTOT))
  # Remove erroneous spaces from Family ID column in Excel and reimport
  SDQ$Pingault_ID <- gsub(gsub('PNG_05A35003UD','',SDQ$Pingault_ID[1]),'', SDQ$Pingault_ID)
  
  # Standardise SDQ scores
  SDQ_standard <- SDQ %>%
    mutate_if(is.numeric, scale)
  
  # Check that we get mean of 0 and SD of 1
  colMeans(subset(SDQ_standard, select=c(BEBDTOT, CEBDTOT, DDDEBDTOT, FEBDTOT)), na.rm=TRUE)
  apply(subset(SDQ_standard, select=c(BEBDTOT, CEBDTOT, DDDEBDTOT, FEBDTOT)), 2, sd, na.rm = TRUE)
  
  # Standardise PGSs
  EApgs <- EApgs %>%
    mutate_if(is.numeric, scale)
  CPpgs <- CPpgs %>%
    mutate_if(is.numeric, scale)
  NCPpgs <- NCPpgs %>%
    mutate_if(is.numeric, scale)
  
  # Check that we get means of 0 and SDs of 1
  mean(EApgs$final_pred_auto, na.rm=TRUE)
  sd(EApgs$final_pred_auto, na.rm = TRUE)
  mean(CPpgs$final_pred_auto, na.rm=TRUE)
  sd(CPpgs$final_pred_auto, na.rm = TRUE)  
  mean(NCPpgs$final_pred_auto, na.rm=TRUE)
  sd(NCPpgs$final_pred_auto, na.rm = TRUE)
  
  # Create dataframes containing mental health scores and PGSs with one family trio per row
  # Educational attainment PGS
  {  
    PGS <- EApgs
    # Merge PGS with data for matching IDs
    matches_PGS <- merge(matches, PGS, by.x="Pingault_ID", by.y="sample.ID")
    matches_PGS[9:10]<-NULL # remove extra parental ID columns
    
    # Create data frame for child information
    child_PGS <- subset(matches_PGS, MFC == 'C         ')  # note dataframe also contains spaces after trio ID
    child_PGS_SDQ <- merge(SDQ_standard, child_PGS, by = "Pingault_ID") # add child total SDQ score
    
    # Rename columns to label as child data (_C)
    child_data_EA <- child_PGS_SDQ %>% 
      rename(
        CM_ID = Pingault_ID,
        SDQ2 = BEBDTOT,
        SDQ3 = CEBDTOT,
        SDQ4 = DDDEBDTOT,
        SDQ6 = FEBDTOT,
        EA_PGS_C = final_pred_auto, 
        SEX_C = SEX,
        REGION_C = REGION,
        ETHNICITY_C = ETHNICITY
      )
    child_data_EA$family.ID=NULL # remove duplicate fam ID column
    
    # create dataframe for mother information
    # subset mother PGS and rename to label as mother data (_M)
    mother_PGS <- subset(matches_PGS, MFC == 'M         ')
    mother_data_EA <- mother_PGS %>% 
      rename(
        M_ID = Pingault_ID,
        EA_PGS_M = final_pred_auto, 
        SEX_M = SEX,
        REGION_M = REGION,
        ETHNICITY_M = ETHNICITY
      )
    mother_data_EA$family.ID=NULL # remove duplicate fam ID column
    
    # create dataframe for father information
    # subset father PGS and rename to label as father data (_F)
    father_PGS <- subset(matches_PGS, MFC == 'F         ')
    father_data_EA <- father_PGS %>% 
      rename(
        F_ID = Pingault_ID,
        EA_PGS_F = final_pred_auto, 
        SEX_F = SEX,
        REGION_F = REGION,
        ETHNICITY_F = ETHNICITY
      )
    father_data_EA$family.ID=NULL # remove duplicate fam ID column
    
    # merge child, mother, father by family ID
    EA_data <- list(child_data_EA, mother_data_EA, father_data_EA) %>% reduce(left_join, by = "Pingault_FID")
  }
  
  # Cognitive performance PGS
  {
    PGS <- CPpgs
    
    # merge PGS with data for matching IDs
    matches_PGS <- merge(matches, PGS, by.x="Pingault_ID", by.y="sample.ID")
    matches_PGS[9:10]<-NULL # remove extra parental ID columns
    
    # create data frame for child information
    child_PGS <- subset(matches_PGS, MFC == 'C         ')
    child_PGS_SDQ <- merge(SDQ_standard, child_PGS, by = "Pingault_ID") 
    
    # Rename columns to label as child data (_C)
    child_data_CP <- child_PGS_SDQ %>% 
      rename(
        CM_ID = Pingault_ID,
        SDQ2 = BEBDTOT,
        SDQ3 = CEBDTOT,
        SDQ4 = DDDEBDTOT,
        SDQ6 = FEBDTOT,
        CP_PGS_C = final_pred_auto, 
        SEX_C = SEX,
        REGION_C = REGION,
        ETHNICITY_C = ETHNICITY
      )
    child_data_CP$family.ID=NULL # remove duplicate fam ID column
    
    # create dataframe for mother information
    # subset mother PGS and rename to label as mother data (_M)
    mother_PGS <- subset(matches_PGS, MFC == 'M         ')
    mother_data_CP <- mother_PGS %>% 
      rename(
        M_ID = Pingault_ID,
        CP_PGS_M = final_pred_auto, 
        SEX_M = SEX,
        REGION_M = REGION,
        ETHNICITY_M = ETHNICITY
      )
    mother_data_CP$family.ID=NULL # remove duplicate fam ID column
    
    # create dataframe for father information
    # subset father PGS and rename to label as father data (_F)
    father_PGS <- subset(matches_PGS, MFC == 'F         ')
    father_data_CP <- father_PGS %>% 
      rename(
        F_ID = Pingault_ID,
        CP_PGS_F = final_pred_auto, 
        SEX_F = SEX,
        REGION_F = REGION,
        ETHNICITY_F = ETHNICITY
      )
    father_data_CP$family.ID=NULL # remove duplicate fam ID column
    
    # merge child, mother, father by family ID
    CP_data <- list(child_data_CP, mother_data_CP, father_data_CP) %>% reduce(left_join, by = "Pingault_FID")
  }
  
  # Non-cognitive PGS
  {
    PGS <- NCPpgs
    # merge PGS with matches data
    matches_PGS <- merge(matches, PGS, by.x="Pingault_ID", by.y="sample.ID")
    matches_PGS[9:10]<-NULL #remove extra parental ID columns
    
    # create data frame for child information
    child_PGS <- subset(matches_PGS, MFC == 'C         ')
    child_PGS_SDQ <- merge(SDQ_standard, child_PGS, by = "Pingault_ID") 
    # Rename columns to label as child data (_C)
    child_data_NCP <- child_PGS_SDQ %>%
      rename(
        CM_ID = Pingault_ID,
        SDQ2 = BEBDTOT,
        SDQ3 = CEBDTOT,
        SDQ4 = DDDEBDTOT,
        SDQ6 = FEBDTOT,
        NCP_PGS_C = final_pred_auto, 
        SEX_C = SEX,
        REGION_C = REGION,
        ETHNICITY_C = ETHNICITY
      )
    child_data_NCP$family.ID=NULL # remove duplicate fam ID column
    
    # create dataframe for mother information
    # subset mother PGS and rename to label as mother data (_M)
    mother_PGS <- subset(matches_PGS, MFC == 'M         ')
    mother_data_NCP <- mother_PGS %>%
      rename(
        M_ID = Pingault_ID,
        NCP_PGS_M = final_pred_auto, 
        SEX_M = SEX,
        REGION_M = REGION,
        ETHNICITY_M = ETHNICITY
      )
    mother_data_NCP$family.ID=NULL # remove duplicate fam ID column
    
    # create dataframe for father information
    # subset father PGS and rename to label as father data (_F)
    father_PGS <- subset(matches_PGS, MFC == 'F         ')
    father_data_NCP <- father_PGS %>%
      rename(
        F_ID = Pingault_ID,
        NCP_PGS_F = final_pred_auto, 
        SEX_F = SEX,
        REGION_F = REGION,
        ETHNICITY_F = ETHNICITY
      )
    father_data_NCP$family.ID=NULL # remove duplicate fam ID column
    
    # merge child, mother, father by family ID
    NCP_data <- list(child_data_NCP, mother_data_NCP, father_data_NCP) %>% reduce(left_join, by = "Pingault_FID")
  }
}
```

### Merge data for each trait into a single data frame.

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}

{
  # merge EA, CP and NCP
  EA_data_subset <- subset(EA_data, select=-c(PNUM.x, MFC.x, PNUM.y, MFC.y, PNUM, MFC))
  CP_data_subset <- subset(CP_data, select=c(Pingault_FID, CP_PGS_C, CP_PGS_M, CP_PGS_F))
  NCP_data_subset <- subset(NCP_data, select=c(Pingault_FID, NCP_PGS_C, NCP_PGS_M, NCP_PGS_F))
  all_data <- list(EA_data_subset, CP_data_subset, NCP_data_subset) %>% reduce(left_join, by = "Pingault_FID")
  
}

```

### Add in population stratification Principle components (PSPCs)

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}

{# load PC file (20PCs)
  PS_PC <- read.table("C:/Users/Jose Morosoli/Documents/UCL/MCS/mcs_imputed_allchr_minQC_updateID_hetQC_relatedQC_PCs.txt", sep = "\t", header=TRUE)
  PSPCID <- subset(PS_PC, select=-c(FID))
  
  # inspect distribution
  colMeans(select(PSPCID, starts_with("PC")))
  apply(select(PSPCID, starts_with("PC")), 2, sd, na.rm = TRUE)
  
  # Match participant IDs to their PSPCs
  trio_ID <- subset(matches, select=c(Pingault_ID, Pingault_FID, MFC))
  trio_ID_PSPC <- merge(trio_ID, PSPCID, by.x="Pingault_ID", by.y="IID")
  
  # select CHILD PCs only - only these will be included in analysis
  child_PSPC <- subset(trio_ID_PSPC, MFC == 'C         ', select=-c(Pingault_ID, MFC))
  child_data_PSPC <- child_PSPC %>% rename_with(.fn = ~paste0(., "_C"), .cols = c(starts_with("PC"))) #add indicator that this is child info
  # scale PCs to mean=0, SD=1
  library(dplyr)
  child_PSPC_scaled <- child_data_PSPC %>%
    mutate_if(is.numeric, scale)
  
  # check scaling worked
  colMeans(select(child_PSPC_scaled, starts_with("PC")))
  apply(select(child_PSPC_scaled, starts_with("PC")), 2, sd, na.rm = TRUE)
  
  # add child PSPCs to the main data frame
  library(tidyverse)
  all_data_PSPC <- list(all_data, child_PSPC_scaled) %>% reduce(left_join, by = "Pingault_FID")
  head(all_data_PSPC)
  colMeans(select(child_data_PSPC, starts_with("PC"))) #double check scaling worked
  apply(select(all_data_PSPC, starts_with("PC")), 2, sd, na.rm = TRUE)
}

# Load MCS data
CMstructure <- read.spss(paste(file_path,"GENDAC_PINGAULT_mcs_cm_structure_2021_05_12.sav",sep = ""), to.data.frame=TRUE)

# Extract internalising and externalising subscales
internal_raw <- subset(CMstructure, select = c(Pingault_ID,BEMOTION,BPEER,CEMOTION,CPEER,DDEMOTION,DDPEER,FEMOTION,FPEER))
external_raw <- subset(CMstructure, select =c(Pingault_ID,BCONDUCT,BHYPER,CCONDUCT,CHYPER,DDCONDUCT,DDHYPER,FCONDUCT,FHYPER))

# Remove erroneous spaces in data using excel, and re-import
internal1 <- internal_raw
external1 <- external_raw
internal1$Pingault_ID <- gsub(gsub('PNG_05A35003UD','',internal_raw$Pingault_ID[1]),'', internal_raw$Pingault_ID)
external1$Pingault_ID <- gsub(gsub('PNG_05A35003UD','',external_raw$Pingault_ID[1]),'', external_raw$Pingault_ID)

# Make factors numeric characters
internal1$BEMOTION <- as.numeric(as.character(internal1$BEMOTION))
internal1$BPEER <- as.numeric(as.character(internal1$BPEER))
internal1$CEMOTION <- as.numeric(as.character(internal1$CEMOTION))
internal1$CPEER <- as.numeric(as.character(internal1$CPEER))
internal1$DDEMOTION <- as.numeric(as.character(internal1$DDEMOTION))
internal1$DDPEER <- as.numeric(as.character(internal1$DDPEER))
internal1$FEMOTION <- as.numeric(as.character(internal1$FEMOTION))
internal1$FPEER <- as.numeric(as.character(internal1$FPEER))
external1$BCONDUCT <- as.numeric(as.character(external1$BCONDUCT))
external1$BHYPER <- as.numeric(as.character(external1$BHYPER))
external1$CCONDUCT <- as.numeric(as.character(external1$CCONDUCT))
external1$CHYPER <- as.numeric(as.character(external1$CHYPER))
external1$DDHYPER <- as.numeric(as.character(external1$DDHYPER))
external1$DDHYPER <- as.numeric(as.character(external1$DDHYPER))
external1$FCONDUCT <- as.numeric(as.character(external1$FCONDUCT))
external1$FHYPER <- as.numeric(as.character(external1$FHYPER))

# Sum subscales to make internalising and externalising factors per sweep
internal1$BINT <- internal1$BEMOTION + internal1$BPEER
internal1$CINT <- internal1$CEMOTION + internal1$CPEER
internal1$DINT <- internal1$DDEMOTION + internal1$DDPEER
internal1$FINT <- internal1$FEMOTION + internal1$FPEER
external1$BEXT <- external1$BCONDUCT + external1$BHYPER
external1$CEXT <- external1$CCONDUCT + external1$CHYPER
external1$DEXT <- external1$DDCONDUCT + external1$DDHYPER
external1$FEXT <- external1$FCONDUCT + external1$FHYPER

# Merge int and ext scores
INTEXT <- merge(internal1, external1, by="Pingault_ID")
myData <- merge(all_data_PSPC, INTEXT, by.x="CM_ID", by.y="Pingault_ID")
myData <- merge(myData, SDQ, by.x="CM_ID", by.y="Pingault_ID")


```

## Standardise phenotipic variables within time point

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}

# Notes:
# Males = 1; Females = 0

myData$zBINT <- scale(myData$BINT, center = T, scale = T)
myData$zCINT <- scale(myData$CINT, center = T, scale = T)
myData$zDINT <- scale(myData$DINT, center = T, scale = T)
myData$zFINT <- scale(myData$FINT, center = T, scale = T)

myData$zBEXT <- scale(myData$BEXT, center = T, scale = T)
myData$zCEXT <- scale(myData$CEXT, center = T, scale = T)
myData$zDEXT <- scale(myData$DEXT, center = T, scale = T)
myData$zFEXT <- scale(myData$FEXT, center = T, scale = T)

```

## Create cross clinical threshold (>=15) variable

Notes: Based on total score (combination of externalising and internalising difficulties)

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}

myData$BCCT <- 0; myData$BCCT[myData$BEBDTOT>=15] <- 1  
myData$CCCT <- 0; myData$CCCT[myData$CEBDTOT>=15] <- 1
myData$DCCT <- 0; myData$DCCT[myData$DDDEBDTOT>=15] <- 1
myData$FCCT <- 0; myData$FCCT[myData$FEBDTOT>=15] <- 1

table(myData$BCCT[myData$SEX==0])/sum(table(myData$BCCT[myData$SEX==0]))*100
table(myData$CCCT[myData$SEX==0])/sum(table(myData$CCCT[myData$SEX==0]))*100
table(myData$DCCT[myData$SEX==0])/sum(table(myData$DCCT[myData$SEX==0]))*100
table(myData$FCCT[myData$SEX==0])/sum(table(myData$FCCT[myData$SEX==0]))*100

table(myData$BCCT[myData$SEX==1])/sum(table(myData$BCCT[myData$SEX==1]))*100
table(myData$CCCT[myData$SEX==1])/sum(table(myData$CCCT[myData$SEX==1]))*100
table(myData$DCCT[myData$SEX==1])/sum(table(myData$DCCT[myData$SEX==1]))*100
table(myData$FCCT[myData$SEX==1])/sum(table(myData$FCCT[myData$SEX==1]))*100
```

### Additional step: Obtain means and standard deviations by sex

```{r echo=T, results='hide', error=F, warning=F, message=F, eval=F}
# Load libraries
library(psych)

describeBy(myData[,c('BEBDTOT','CEBDTOT','DDDEBDTOT','FEBDTOT')], group = myData$SEX)
describeBy(myData[,c('BEXT','CEXT','DEXT','FEXT')], group = myData$SEX)
describeBy(myData[,c('BINT','CINT','DINT','FINT')], group = myData$SEX)

```

